---
title: "Linear Projection as Optimization Problem"
author: "Hans W. Borchers"
date: "2018-04-11"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
  pdf_document: 
    keep_tex: yes
---

## Introduction

### Linear Projection as Optimization Task

Given an affine subspace $S$ of $\mathrm{R}^n$, defined through a matrix equation $A x = b$ with a matrix $A$ of size $(m, n)$ and a vector $b$ of length $m$, the task is to find the *projection* of a point $p_0$ onto this subspace.
We could solve this problem as a pure Linear Algebra task. Then we would look at a decomposition of $\mathrm{R}^n$ as an orthogonal sum $A \oplus U$ of $A$ and another subspace $U$. Then represent $p_0$ in a basis of $U$.

Here we will solve the problem as an optimization task. If $p$ is the projection of $p0$ on $A$, then the distance between $p0$ and $p$ is minimal for all points in $S$, that is $p$ with $A p = b$. We will solve this for an example with $n = 1000$ and $m = 100$.

### Generating Example Data

```{r}
set.seed(65537)
n = 1000; m = 100
p0 <- rep(0, n)
Aeq <- matrix(runif(m*n), nrow = m)
beq <- rep(1, m)
```


## Solving the Optimization Task

### Solution with *quadprog*

```{r}
library(quadprog)

Q = diag(n)
d = as.matrix(p0)
A = t(Aeq)
b = as.matrix(beq)

system.time( sol <- solve.QP(Q, d, A, b, meq=m) )
```
```{r}
p = sol$solution
d = sqrt(sum((p - p0)^2))
cat("The minimal distance is:", d)
```


### Solution with *CVXR*

```{r}
library(CVXR)

x   = Variable(n)
obj = Minimize( sum((p0 - x)^2) )
con = list(Aeq %*% x == beq)
pro = Problem(obj, con)

system.time( sol <- solve(pro) )

cat("The minimal distance is:", sqrt(sol$value))
```

We can also try out the other available solver, SCS:

```{r}
system.time( sol <- solve(pro, solver = "SCS") )
```


### Use the ECOS Solver Directly

Instead of employing *CVXR* we can call the underlying ECOS solver directly. First we define the `mindist` function that builds the necessary matrices and then sends them to `ECOS_csolve`.

```{r}
library(ECOSolveR)

mindist <- function(x0, Aeq, beq) {
    n <- length(x0)
    cc <- c(rep(0, n), 1)
    A_eq <- Matrix::Matrix(cbind(Aeq, 0), sparse = TRUE)
    b_eq <- beq
    A2 <- rbind(c(rep(0, n), -1),
                cbind(diag(1, n), 0))
    G <- Matrix::Matrix(rbind(A2), sparse = TRUE)
    h <- c(0, x0)
    sol <- ECOS_csolve(cc, G = G, h = h,
                       dims = list(q = as.integer(n+1)),
                       A = A_eq, b = b_eq)
    return(list(point = sol$x[1:n], dist = sol$x[n+1],
                message = sol$infostring))
}
```

```{r}
system.time(sol <- mindist(p0, Aeq, beq))

cat("The minimal distance is:", sol$dist)
```

### Prepare for Several Projections

Or we set up the matrices needed before we call the solver:

```{r}
n = length(p0)
cc = c(rep(0, n), 1)
A_eq = Matrix::Matrix(cbind(Aeq, 0), sparse = TRUE)
b_eq = beq
A2 = rbind(c(rep(0, n), -1),
           cbind(diag(1, n), 0))
G <- Matrix::Matrix(rbind(A2), sparse = TRUE)
h <- c(0, p0)
system.time(
    sol <- ECOS_csolve(cc, G = G, h = h,
                       dims = list(q = as.integer(n+1)),
                       A = A_eq, b = b_eq)
)
dist = sol$x[n+1]
cat("The minimal distance is:", d)
```

